---
title: "Example Notebook"
author: "Adrien Osakwe"
output: html_notebook
---

Notebook containing analysis and results for the Iris dataset project.

## Iris Dataset

The iris dataset is composed of 50 flowers from three different species. Measurements include width and length (centimeters) for petals and sepals of each flower.

### Feature visualization

```{r}
mini_iris <- iris %>%
  filter(Species != 'versicolor')

ggplot(mini_iris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) +
  geom_point(size = 3, alpha = 0.7) +  # Add points with some transparency
  theme_minimal() +  # Use a clean theme
  labs(
    title = "Iris Dataset: Sepal Dimensions by Species",
    x = "Sepal Length",
    y = "Sepal Width"
  )

```

## Objective

Goal of this analysis is to identify features capable of distinguishing flower species with high accuracy. We achieve this task using the **logistic regression model**, focusing on 1 vs. all comparisons.

### Logistic Regression

From a probabilistic perspective, the logistic regression follows a Bernoulli distribution

$$
y_i \sim Bernouilli(\pi_i)
$$

Where $y_i$ is the classification for observation $i$. $\pi_i$ represents the Bernoulli parameter for observation $i$ and is calculated as

$$
\pi_i = \sigma(x\beta) = P(y_i = 1| x)=  \frac{1}{1 + \exp(-x\beta)}
$$

Logistic regression extends the linear model by using the sigmoid/logistic function $\sigma(x\beta)$ which allows it to bound its outputs between 0 and 1.

## Model Fitting

```{r}


iris.fit <- glm(Species ~ Sepal.Length +  Sepal.Width,
                family = 'binomial',
                data = mini_iris)

mini_iris$Prob <- predict.glm(iris.fit,
                              mini_iris[,c('Sepal.Length','Sepal.Width')],
                              type = 'response')

mini_iris$Pred <- ifelse(mini_iris$Prob >= 0.5,"virginica", "setosa")

mini_iris$Result <- mini_iris$Species == mini_iris$Pred

```

## Model Visualization

```{r}

ggplot(mini_iris, aes(x = Sepal.Length, y = Sepal.Width, color = Pred)) +
  geom_point(size = 3, alpha = 0.7) +  # Add points with some transparency
  theme_minimal() +  # Use a clean theme
  labs(
    title = "Iris Dataset: Training Classification Performance",
    x = "Sepal Length",
    y = "Sepal Width"
  )


mini_iris$Num_Species <- ifelse(mini_iris$Species == 'setosa', 0,1)
iris.lm <- lm(Num_Species ~ Sepal.Length +  Sepal.Width,
              data = mini_iris)

mini_iris$Num_pred <- predict.lm(iris.lm,mini_iris)

summary(mini_iris$Num_pred)

summary(mini_iris$Prob)


```

### Part II

```{r}

mini_iris <- iris %>%
  filter(Species != 'setosa')
ggplot(mini_iris, aes(x = Petal.Length, y = Petal.Width, color = Species)) +
  geom_point(size = 3, alpha = 0.7) +  # Add points with some transparency
  theme_minimal() +  # Use a clean theme
  labs(
    title = "Iris Dataset: Dimensions by Species",
    x = "Petal Length",
    y = "Petal Width"
  )

iris.fit <- glm(Species ~ Sepal.Length + Sepal.Width,
                family = 'binomial',
                data = mini_iris)

```

### Part III

```{r}

iris_pred <- predict.glm(iris.fit,
                         mini_iris[,c('Sepal.Length','Sepal.Width')],
                         type = 'response')
as.data.frame(table(mini_iris$Species == ifelse(iris_pred >= 0.5, 'virginica','versicolor')))
```
